### 2021-10-25 SujetStage/2021-pNetsubjectM2.tex
version v0

\begin{document}
\begin{center}
\resizebox{0.5\textwidth}{!}{\bf Internship subject (Master 2 or Master 1)}
\\\bigskip
\resizebox{\textwidth}{!}{\bf Refinement for open automata}
%  Systolisation d'un réseau de processus polyédrique}
\end{center}
\medskip

\begin{description}
\item[\bf Main advisor:] Rabéa Ameur-Boulifa 
\item[\bf Co-advisors:] Ludovic Henrio and Eric Madelaine
\item[\bf Place:] Eurecom  -- Sophia Antipolis
\\
or\\
Laboratoire de l'Informatique du Parallélisme (LIP) --
  \'Ecole Normale Supérieure de Lyon
\end{description}

@@

Establishing equivalences or refinement relations between programs or system is crucial  for verifying correctness of programs, by establishing that one implementation is the refinement of a specification.

In the previous years, we have studied theoretical foundations for open
systems and our formalism, called open automata, is able to represent operators of composition of processes,  represented
as hierarchically composed automata with holes and parameters. 
Our long 
term goal is to 
develop a methodology combining symbolic operational semantic and bisimulation 
equivalences with deductive reasoning on the data part, 
and in practice combining bisimulation algorithms with SMT solvers to get automatic 
procedures proving equational properties
of these open systems. In the last years, we designed a weak bisimulation theory for  open automata and a translation from a specification language that we used, called pNets, to open automata~\cite{arxiv-weakbisim,henrio:Forte2016,hou:hal-02406098}.

Among the properties of our formalisms, we are interested in compositionality: If two systems are proven equivalent they will be undistinguishable by their context, and they will also be undistinguishable when their holes are filled with equivalent systems.
The article \cite{arxiv-weakbisim} is illustrated with a transport protocol running example; it shows the characteristics of our formalism and our bisimulation relations.

@@

The main objective of the internship is to study refinement theory for open automata.  
Our purpose is to define a refinement relation between the symbolic models allowing refinement verification for a class of open systems.

The internship will start with a study of the semantics of a simple language called value passing CCS. The first objective of this internship is to design a semantics value passing CCS in terms of open automata. Especially, we will be able to extend the semantics by taking into account the semantics of operators for composing CCS processes.

After this preliminary step, the next objective will be to design a refinement theory for open automata. We will make sure that this relation is specified in a constructive manner, so that an algorithm could be derived from the specification in the future~\cite{Bellegarde:MEMOCODE2003}.
The relationship between refinement and automata composition should also play a major role in this definition.

%The purpose is to define the refinement semantics as a relation between open automata that can be classified in  and we will make sure that an algorithm could be designed to verify refinement can also be verified for finite state systems using an algorithmic method.

The next steps in the internship will rely on these two initial works; they will consist in addressing some of the following (independent) objectives:
\begin{itemize}
\item Express and prove compositionality properties for the refinement relation over open automata.
\item Study formally the behaviour of CCS operators wrt refinement.
\item Compare this theory with the state of the art of the refinement semantics.
\item Design the algorithm that checks refinement between two open automata.
\end{itemize}

@@

\bibitem{Bellegarde:MEMOCODE2003}
Fran{\c{c}}oise Bellegarde, Celina Charlet and Olga Kouchnarenko:
\newblock {How to Compute the Refinement Relation for Parameterized Systems}.
\newblock 1st {ACM} {\&} {IEEE} International Conference on Formal Methods and Models for Co-Design {(MEMOCODE} 2003), 24-26 June 2003, Mont Saint-Michel, France. 
\newblock DOI: 10.1109/MEMCOD.2003.1210095

>>>>

\begin{document}
\begin{center}
\resizebox{0.5\textwidth}{!}{\bf Internship subject (Master 2 or Master 1)}
\\\bigskip
\resizebox{\textwidth}{!}{\bf Session Types meet Open Automata}
%  Systolisation d'un réseau de processus polyédrique}
\end{center}
\medskip

\begin{description}
\item[\bf Main advisor:] Rabéa Ameur-Boulifa ?
\item[\bf Co-advisors:] Ludovic Henrio and Eric Madelaine?
\item[\bf Place:] Eurecom  -- Sophia Antipolis
\\
or\\
Laboratoire de l'Informatique du Parallélisme (LIP) --
  \'Ecole Normale Supérieure de Lyon
\end{description}

@@

In the previous years, we have studied theoretical foundations for open
systems and we defined  open automata \cite{arxiv-weakbisim,henrio:Forte2016,hou:hal-02406098} that can be seen as labelled transition systems (LTSs) with parameters and holes. The transitions of open
automata are much more complex than transitions of an LTS as the ﬁring on a transition depends on parameters and actions that are symbolic.

Establishing  whether two open automata are composable or not require the designer to provide the behaviour of all holes in a context where they are used; which consists in describing all the interactions and the series of inputs and outputs of data exchanged  between the automata.  To make this task reusable by other designers, we are interested in characterising  the behavioural condition that must be satisfied by a process that would be plugged into a hole to satisfy the composability condition.

To tackle this issue, we ask: Can we synthesise a global type from a collection of automata so as to ensure the composability of an open automaton whose hole is compatible with this global type?

@@

The main objective of the internship  deals with the characterization of dynamic execution of open automata as behavioural types. Our purpose is to define the composability condition of two open automata.

The internship will start with a study of the semantics of  open automata and the notion behavioural types \cite{Hans:2016}, in particular the notion of session types that allows to describe the dynamic aspects of processes.

@@

\bibitem{Hans:2016}
Hans H\"uttel, Ivan Lanese, Vasco T. Vasconcelos, Lu\`is Caires, Marco Carbone, Pierre-Malo Deni\'elou, Dimitris Mostrous, Luca Padovani, Ant\`onio Ravara, Emilio Tuosto, Hugo Torres Vieira, Gianluigi Zavattaro:
\newblock {Foundations of Session Types and Behavioural Contracts}.
\newblock ACM Comput. Surv. 49(1): 3:1-3:36 (2016)
